---
title: "Data Processing"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
    df_print: kable
---

## Overview

This Rmarkdown report processes and cleans the Eurovision Song Contest (ESC) data for modeling. 

This includes:

1. Redefine variables as factor or numeric
2. Dividing the variables into the three predefined groups; 
    * Performance
    * External 
    * Competition
3. Normalize the Numeric Data to have mean 0 and standard deviation 1
4. Dummy Encoding all Categorical Factor levels.
5. Data Reduction
    * Redundant Variables
    * Variables of Linear Combinations
    * Categorical Variables via Chi-Squared Tests of Association

```{r preliminaries, message = FALSE, echo = FALSE}
# load required libraries
library(rmarkdown)
library(knitr)
library(ggplot2)
library(dplyr)
# load in custom utility functions
source("../scripts/utilities/column_to_factor.R")
source("../scripts/utilities/TC_FC_missing_observations.R")
source("../scripts/utilities/plot_missings_effect.R")
source("../scripts/utilities/extract_numeric_data.R")
source("../scripts/utilities/extract_factor_data.R")
source("../scripts/utilities/categorical_dummy_encoding.R")
source("../scripts/utilities/range_standardise_data.R")
source("../scripts/utilities/normalise_data.R")
source("../scripts/utilities/range_normalisation.R")
source("../scripts/utilities/data_normalisation.R")
source("../scripts/utilities/data_reduction_chisq.R")
source("../scripts/utilities/data_reduction_corr.R")
```

```{r data-load, echo = FALSE}
# load in the raw ESC 2016 data for the analysis
ESCdata <- read.csv(file = "../data/ESC_2016_voting_data.csv", header = T)
# order the data by To_country, From_country, Round and OOA
ESCdata <- ESCdata %>% arrange(To_country, From_country, Round, OOA)
# drop the id variable
ESCdata <- ESCdata %>% select(-c(id))
```

## Redefine Factor Variables 

Some of the numeric music features need to be redefined as nominal variables, the variables are key, mode and time signature.

```{r redefine-factor-variables, echo = FALSE}
# define the columns to be converted to factor variables
to_factor_cols <- c('key', 'mode', 'time_signature', 'VBlocs1_FC', 'VBlocs2_FC', 'VBlocs1_TC', 'VBlocs2_TC')
# call the column to factor function 
ESCdata <- column_to_factor(dataset = ESCdata, col_names = to_factor_cols)
# print post types
initial_data_types <- as.data.frame(sapply(ESCdata, class))
colnames(initial_data_types) <- c("Data Type")
initial_data_types
```

## Remove Missing Observations

```{r remove-missing-observations-stats, echo = FALSE}
# NA values per column
na_count_per_column <- sapply(ESCdata, function(y) sum(length(which(is.na(y)))) / nrow(ESCdata))
# write to a csv file
write.csv(na_count_per_column, '../report/stats/NA_prop_per_columns.csv')
# filter our rows with missing data values
completedata <- ESCdata %>% na.omit

# extract out unique from countries in order
from_countries <- unique(ESCdata$From_countr) %>% sort()
# generate count stats for the from countries before and after missings are removed
TC_FC_missing_obseration_df <- TC_FC_missing_observations(from_countries = from_countries, orig_data = ESCdata, comp_data = completedata)
# write the TC_FC_missing_observations_df to a csv file
write.csv(x = TC_FC_missing_obseration_df, file = "../report/stats/TC_FC_missing_obseration_df.csv")
```

There is a total of `r nrow(ESCdata) - nrow(ESCdata %>% na.omit)` rows missing from the data. This corresponds to approximately `r round((nrow(ESCdata) - nrow(ESCdata %>% na.omit)) / nrow(ESCdata) * 100, 2)`% of the data. Removing all of these rows with missing values leaves a total of `r nrow(completedata)` rows remaining. This is a substantial loss of data, a notable limitation to the research, and an area of improvement for future research iterations. An alternative solution could be to use different sources such as the world bank for the migration / diaspora based data. However at time of research, these data repositories did not store migration / diaspora information by country of birth / citizenship.

http://www.worldbank.org/en/topic/migrationremittancesdiasporaissues/brief/migration-remittances-data

```{r remove-missing-observations-plots, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}
# The Effect on To_country, From_country, Points
voting_factors <- c('To_country', 'From_country', 'Points')
# call the plot missing function
plot_missings_effect(col_name = voting_factors, orig_data = ESCdata, comp_data = completedata, output_dir = '../report/plots/missings/')
```

## Generate Factor Blocs

```{r voting-factors, echo = FALSE}
# extract all the voting factors
voting_factors <- completedata %>% subset(select = voting_factors)
```

```{r competition-factors, echo = FALSE}
# define the competition factor column names
comp_factors_names <- c('Average_Points', 'Round', 'Voting_Method', 'Host_Nation', 'OOA')
# extract the competition factor column names
competition_factors <- completedata %>% subset(select = comp_factors_names)
```

```{r external-factors, echo = FALSE}
# define the external factor column names
ext_factors_names <- c('VBlocs1_FC', 'VBlocs2_FC', 'VBlocs1_TC', 'VBlocs2_TC', 'ComVBlocs1', 'ComVBlocs2', 
                      'FC_LANGFAM', 'TC_LANGFAM', 'ComLANGFAM', 'Neighbours', 'TC_NumNeigh', 'FC_NonCOB',
                      'FC_NonCitzens', 'FC_COB', 'FC_Citizens', 'FC_Population', 'METRIC_COB', 'METRIC_Citizens',
                      'METRIC_COBCit', 'FC_GDP_mil', 'TC_GDP_mil', 'GDP_PROP', 'FC_CAP_LAT', 'FC_CAP_LON', 
                      'TC_CAP_LAT', 'TC_CAP_LON', 'CAP_DIST_km')
# extract the external factor column names
external_factors <- completedata %>% subset(select = ext_factors_names)
```

```{r performance-factors, echo = FALSE}
# define the performance factor column names
perf_factors_names <- c('TC_PerfType', 'TC_SingerGender', 'FC_SONGLANG', 'TC_SONGLANG', 'ComSONGLAN', 
                       'danceability', 'energy', 'key', 'loudiness', 'mode', 'speechiness', 'acousticness',
                       'instrumentalness', 'liveness', 'valence', 'tempo', 'duration_ms', 'time_signature')
# extract the performance factor column names
performance_factors <- completedata %>% subset(select = perf_factors_names)
```

In total, there are:

* `r ncol(voting_factors)` Voting Factors
* `r ncol(competition_factors)` Competition Factors
* `r ncol(external_factors)` External Factors
* `r ncol(performance_factors)` Performance Factors

## Extract Numeric & Categorical Features

```{r competition-features, echo = FALSE}
# (i) Numeric Variables
competition_factors_numeric <- extract_numeric_data(dataset = competition_factors)
# (ii) Categorical competition Variables
competition_factors_categorical <- extract_factor_data(dataset = competition_factors)
```

```{r external-features, echo = FALSE}
# (i) Numeric Variables
external_factors_numeric <- extract_numeric_data(dataset = external_factors)
# (ii) Categorical Variables
external_factors_categorical <- extract_factor_data(dataset = external_factors)
```

```{r performance-features, echo = FALSE}
# (i) Numeric Variables
performance_factors_numeric <- extract_numeric_data(dataset = performance_factors)
# (ii) Categorical Variables
performance_factors_categorical <- extract_factor_data(dataset = performance_factors)
```

```{r voting-features, echo = FALSE}
# (i) Numeric Variables
voting_factors_numeric <- extract_numeric_data(dataset = voting_factors)
#  (ii) Categorical Variables
voting_factors_categorical <- extract_factor_data(dataset = voting_factors)
```

In total, there are:

* `r ncol(competition_factors_numeric)` Numeric Voting Factors
* `r ncol(competition_factors_categorical)` Categorical Voting Factors
* `r ncol(external_factors_numeric)` Numeric Competition Factors
* `r ncol(external_factors_categorical)` Categorical Competition Factors
* `r ncol(performance_factors_numeric)` Numeric External Factors
* `r ncol(performance_factors_categorical)` Categorical External Factors
* `r ncol(voting_factors_numeric)` Numeric Performance Factors
* `r ncol(voting_factors_categorical)` Categorical Performance Factors

## Dummy Encoding Categorical Variables

All categorical variables for each variable bloc are dummy encoded. It is not necessary to dummy encode the voting factors To_country and From_country as we will not be incorporated as model predictor variables

```{r dummy-encoding, echo = FALSE}
# Dummy encode competition factors
competition_factors_categorical <- categorical_dummy_encoding(dataset = competition_factors_categorical)
# Dummy encode external factors
external_factors_categorical <- categorical_dummy_encoding(dataset = external_factors_categorical)
# Dummy encode performance factors
performance_factors_categorical <- categorical_dummy_encoding(dataset = performance_factors_categorical)
```

## Standardise Numeric Data

This section normalizes or range standardizes the numeric variables in each block. There is also a special case for OOA in the competition Bloc. OOA will need to be standardized in relation to each round. This shall be done after the numeric variables from each bloc have been standardized

```{r standardise-competition, echo = FALSE}
# (1) Competition Factors
competition_factors_numeric <- normalise_data(dataset = competition_factors_numeric)
# Check that each column has mean 0
competition_factors_numeric_mean <- round(apply(X = competition_factors_numeric, MARGIN = 2, FUN = mean), digits = 6)
# Check that each column has standard deviation 1
competition_factors_numeric_sd <- round(apply(X = competition_factors_numeric, MARGIN = 2, FUN = sd), digits = 6)
# column bind stats
competition_factors_numeric_stats <- as.data.frame(cbind(competition_factors_numeric_mean, competition_factors_numeric_sd))
colnames(competition_factors_numeric_stats) <- c("mean", "sdev")
competition_factors_numeric_stats
```

```{r standardise-external, echo = FALSE}
# (2) External Factors
external_factors_numeric <- normalise_data(dataset = external_factors_numeric)
# Check that each column has mean 0
external_factors_numeric_mean <- round(apply(X = external_factors_numeric, MARGIN = 2, FUN = mean), digits = 6)
# Check that each column has standard deviation 1
competition_factors_numeric_sd <- round(apply(X = external_factors_numeric, MARGIN = 2, FUN = sd), digits = 6)
# column bind stats
external_factors_numeric_stats <- as.data.frame(cbind(external_factors_numeric_mean, competition_factors_numeric_sd))
colnames(external_factors_numeric_stats) <- c("mean", "sdev")
external_factors_numeric_stats
```

```{r standardise-performance, echo = FALSE}
# (3) Performance Factors
performance_factors_numeric <- normalise_data(dataset = performance_factors_numeric)
# Check that each column has mean 0
performance_factors_numeric_mean <- round(apply(X = performance_factors_numeric, MARGIN = 2, FUN = mean), digits = 6)
# Check that each column has standard deviation 1
performance_factors_numeric_sd <- round(apply(X = performance_factors_numeric, MARGIN = 2, FUN = sd), digits = 6)
# column bind stats
performance_factors_numeric_stats <- as.data.frame(cbind(performance_factors_numeric_mean, performance_factors_numeric_sd))
colnames(performance_factors_numeric_stats) <- c("mean", "sdev")
performance_factors_numeric_stats
```

```{r standardise-ooa, echo = FALSE}
#-- OOA --#

# Extract the relevant features
OOA_df <- completedata %>% subset(select = c('From_country', 'To_country', 'Round', 'OOA'))
# Order the Features
OOA_df <- OOA_df %>% arrange(Round, OOA, To_country, From_country)
# check head of results
head(OOA_df)
# Apply the range standardization
OOA_df[OOA_df$Round == "f", 4] <- range_normalisation(dataset = OOA_df[OOA_df$Round == "f", 4], lb = 0, ub = 1)
OOA_df[OOA_df$Round == "sf1", 4] <- range_normalisation(dataset = OOA_df[OOA_df$Round == "sf1", 4], lb = 0, ub = 1)
OOA_df[OOA_df$Round == "sf2", 4] <- range_normalisation(dataset = OOA_df[OOA_df$Round == "sf2", 4], lb = 0, ub = 1)
head(OOA_df)
# Re-Order the data frame
OOA_df <- OOA_df %>% arrange(To_country, From_country, Round, OOA)
# check the results
head(OOA_df)
# compare with original
head(completedata %>% subset(select = c('From_country', 'To_country', 'Round', 'OOA')))
# check target to overwrite
head(competition_factors_numeric)
# Overwrite the OOA in competition_factors_numeric with the new standardized OOA
competition_factors_numeric[,"OOA"] <- OOA_df[,"OOA"]
```


## Data Reduction

This section performs data reduction whereby Categorical / Numeric variables are removed if:
1. They have only a single type of observation, 0 or 1
2. They form part of a linear combination with other variables
3. They are strongly associated / correlated with other variables

```{r drop-redundant-external, echo = FALSE}
# There are 13 categorical variables that have only 0 observations
zero_ext_fact_cats <- apply(X = external_factors_categorical, MARGIN = 2, FUN = sum) == 0
# extract out the columns with all zero values
zero_ext_fact_cats_cols <- names(which(zero_ext_fact_cats))
# find the columns which do not have all zero values
non_zero_ext_fact_cols <- colnames(external_factors_categorical)[!(colnames(external_factors_categorical) %in% zero_ext_fact_cats_cols)]
# determine how many columns have zero values
n_zero_ext_fact_cats_cols <- length(zero_ext_fact_cats_cols)
# extract the external categorical features
external_factors_categorical <- external_factors_categorical %>% subset(select = non_zero_ext_fact_cols)
```

```{r drop-redundant-competition, echo = FALSE}
# There are no competition categorical variables with only 0 observations
n_competition_factors_categorical <- sum(apply(X = competition_factors_categorical, MARGIN = 2, FUN = sum) == 0)

```

```{r drop-redundant-performance, echo = FALSE}
# There are 0 performance categorical variables with only 0 observations
n_performance_factors_categorical <- sum(apply(X = performance_factors_categorical,  MARGIN = 2, FUN = sum) == 0)
```

In total, there are:

* `r n_zero_ext_fact_cats_cols` Redundant External Categorical Factors
* `r n_competition_factors_categorical` Redundant Competition Factors
* `r n_performance_factors_categorical` Redundant Performance Factors

If two variables are feature a lot of 0s or 1s then there will be a strong association between the two variables as they share a lot of common observations. In such cases, one of the variables can be removed as they both measure the same entity. This lowers the chance of collinearity and reduces the number of dimensions

```{r drop-chisq-external, echo = FALSE}
# extract out the relevant voting blocks for each county
FC_VBlocs1_facts <- external_factors_categorical %>% select(starts_with('VBlocs1_FC')) %>% colnames()
FC_VBlocs2_facts <- external_factors_categorical %>% select(starts_with('VBlocs2_FC')) %>% colnames()
TC_VBlocs1_facts <- external_factors_categorical %>% select(starts_with('VBlocs1_TC')) %>% colnames()
TC_VBlocs2_facts <- external_factors_categorical %>% select(starts_with('VBlocs2_TC')) %>% colnames()
FC_LANGFAM_facts <- external_factors_categorical %>% select(starts_with('FC_LANGFAM')) %>% colnames()
TC_LANGFAM_facts <- external_factors_categorical %>% select(starts_with('TC_LANGFAM')) %>% colnames()
# extract out the column names for significant columns
chisq_tests_FC_VBlocs1_cols <- data_reduction_chisq(dataset = external_factors_categorical, col_names = FC_VBlocs1_facts)
chisq_tests_FC_VBlocs2_cols <- data_reduction_chisq(dataset = external_factors_categorical, col_names = FC_VBlocs2_facts)
chisq_tests_TC_VBlocs1_cols <- data_reduction_chisq(dataset = external_factors_categorical, col_names = TC_VBlocs1_facts)
chisq_tests_TC_VBlocs2_cols <- data_reduction_chisq(dataset = external_factors_categorical, col_names = TC_VBlocs2_facts)
chisq_tests_FC_LANGFAM_cols <- data_reduction_chisq(dataset = external_factors_categorical, col_names = FC_LANGFAM_facts)
chisq_tests_TC_LANGFAM_cols <- data_reduction_chisq(dataset = external_factors_categorical, col_names = TC_LANGFAM_facts)
# concatenate all columns to be removed
remove_vblocs_cols <- c(chisq_tests_FC_VBlocs1_cols, chisq_tests_FC_VBlocs2_cols, chisq_tests_TC_VBlocs1_cols, chisq_tests_TC_VBlocs2_cols)
remove_langfam_cols <- c(chisq_tests_FC_LANGFAM_cols, chisq_tests_TC_LANGFAM_cols)
remove_ext_cols <- c(remove_vblocs_cols, remove_langfam_cols)
# extract all the external factors categorical variables
external_factors_categorical_cols <- colnames(external_factors_categorical)
# determine the columns to keep
keep_ext_cols <- external_factors_categorical_cols[!(external_factors_categorical_cols %in% remove_ext_cols)]
# Remove Unnecessary Categorical Variables 
external_factors_categorical <- external_factors_categorical %>% subset(select = keep_ext_cols)
```

```{r drop-chisq-performance, echo = FALSE}
# extract out the relevant from country and to country song language
FC_SONGLANG_facts <- performance_factors_categorical %>% select(starts_with('FC_SONGLANG')) %>% colnames()
TC_SONGLANG_facts <- performance_factors_categorical %>% select(starts_with('TC_SONGLANG')) %>% colnames()
keys_fact <- performance_factors_categorical %>% select(starts_with('key_')) %>% colnames()
# extract out the column names for significant columns
chisq_tests_FC_SONGLANG_cols <- data_reduction_chisq(dataset = performance_factors_categorical, col_names = FC_SONGLANG_facts)
chisq_tests_TC_SONGLANG_cols <- data_reduction_chisq(dataset = performance_factors_categorical, col_names = TC_SONGLANG_facts)
chisq_tests_keys_cols <- data_reduction_chisq(dataset = performance_factors_categorical, col_names = keys_fact)
# concatenate all columns to be removed
remove_songlang_cols <- c(chisq_tests_FC_SONGLANG_cols, chisq_tests_TC_SONGLANG_cols)
remove_perf_cols <- c(remove_songlang_cols, chisq_tests_keys_cols)
# extract all the external factors categorical variables
performance_factors_categorical_cols <- colnames(performance_factors_categorical)
# determine the columns to keep
keep_perf_cols <- performance_factors_categorical_cols[!(performance_factors_categorical_cols %in% remove_perf_cols)]
# Remove Unnecessary Categorical Variables 
performance_factors_categorical <- performance_factors_categorical %>% subset(select = keep_perf_cols)
```

If two numeric variables are very highly correlated and represent the same entity then it is unnecessary to include them in the data modeling stage. This is particular the case for the migration data. The correlation test function implement is the same as the one that was used during the exploratory analysis section.

```{r drop-corr-migration, echo = FALSE}
# Perform the correlation tests
mig_nums = c('FC_NonCOB', 'FC_NonCitzens', 'FC_COB', 'FC_Citizens', 'FC_Population', 'METRIC_COB','METRIC_Citizens', 'METRIC_COBCit')
# extract out the column names for significant columns
corr_tests_mig_cols <- data_reduction_corr(dataset = external_factors_numeric, col_names = mig_nums)
# extract all the external factors categorical variables
external_factors_numeric_cols <- colnames(external_factors_numeric)
# determine the columns to keep
keep_ext_num_cols <- external_factors_numeric_cols[!(external_factors_numeric_cols %in% corr_tests_mig_cols)]
# Remove Unnecessary Categorical Variables 
external_factors_numeric <- external_factors_numeric %>% subset(select = keep_ext_num_cols)
```

## Data Output

```{r data-output, echo = FALSE}
# combine the various factor blocks together
processed_voting_factors <- cbind(voting_factors_categorical, voting_factors_numeric)
processed_competition_factors <- cbind(competition_factors_categorical, competition_factors_numeric)
processed_external_factors <- cbind(external_factors_categorical, external_factors_numeric)
processed_performance_factors <- cbind(performance_factors_categorical, performance_factors_numeric)
# combine and create the final processed data set
processed_data <- cbind(processed_voting_factors, processed_competition_factors, processed_external_factors, processed_performance_factors)
# write the processed data to a csv file
write.csv(processed_data, '../data/processed_data.csv', row.names = F)
```
