---
title: "Data Modeling"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
    df_print: kable
---

## Overview

### Analysis Methodology

A stratified analysis method will be implemented:

1. Test Overall Data
    * Iteratively create model step wise, 
2. Split and Test Data by Voting Method
    * Can't stratify the data by each country due to a lack of data
    * Need a minimum of 10/20 observations per covariate for regression analysis
    * Split data by voting method, research televote shows more bias than the jury

Note:

* forward and step wise fitting will be utilized using AIC to determine model of best fit
* the models will be evaluated using the car package

### Model Assumptions

Multiple Linear regression (MLR) requires the model residuals to be ~ IID N(0, sigma^2). The model residuals will be standardized for these assessments.

1. Normality Assumptions will be accessed using:
    * Normality tests from the nortest package
    * Visualizations such as histograms, QQ-plots, Residual Plots and Add Variable Plots
2. Constant Variance will be accessed using:
    * non-constant variance test
3. Multi-collinearity will be accessed using:
    * variance inflation factors
4. Outliers will be accessed using:
    * Cooks Distance

```{r preliminaries, message = FALSE, echo = FALSE}
# load the relevant libraries
library(rmarkdown)
library(knitr)
library(car)
library(dplyr)
# load custom functions
source('../scripts/utilities/step_lm_model.R')
```

```{r data-load, echo = FALSE}
#-- Data --#

# load in the historic voting data for deriving the voting blocs
processed_data <- read.csv(file = "../data/processed_data.csv", header = T)
# split the televote data
televote_data <- processed_data %>% filter(Voting_Method_J == 0)
# split out the jury vote data
jury_data <- processed_data %>% filter(Voting_Method_J == 1)
# define the competition factors
competition_factors <- extract_preds_by_cats(cat = 'competition')
# define the performance factors
performance_factors <- extract_preds_by_cats(cat = 'performance')
# define the external factors
external_factors <- extract_preds_by_cats(cat = 'external')
```

\newpage

## Multiple Linear Regression Models

### Overall

```{r model-overall-competition, echo = FALSE}
# fit step wise linear model
overall_competition_model <- step_lm_model(dataset = processed_data, pred_cols = competition_factors)
# extract out significant predictors
overall_sign_competition_factors <- overall_competition_model$sign_factors
```

```{r model-overall-performance, echo = FALSE}
# fit step wise linear model
overall_performance_model <- step_lm_model(dataset = processed_data, pred_cols = performance_factors)
# extract out significant predictors
overall_sign_performance_factors <- overall_performance_model$sign_factors
```

```{r model-overall-external, echo = FALSE}
# fit step wise linear model
overall_external_model <- step_lm_model(dataset = processed_data, pred_cols = external_factors)
# extract out significant predictors
overall_sign_external_factors <- overall_external_model$sign_factors
```

```{r model-overall-all, echo = FALSE}
# define the significant factors
overall_all_factors <- c(overall_sign_competition_factors, overall_sign_performance_factors, overall_sign_external_factors)
# fit step wise linear model
overall_all_model <- step_lm_model(dataset = processed_data, pred_cols = overall_all_factors)
# extract out significant predictors
overall_sign_all_factors <- overall_all_model$sign_factors
```

```{r model-overall-final, echo = FALSE}
# create a model formula
overall_final_model_form <- as.formula(paste('Points ~', paste(overall_sign_all_factors, collapse = ' + ')))
# create the maximum model
overall_final_model <- lm(overall_final_model_form, data = processed_data)
# get model summary
summary(overall_final_model)
# check variance inflation factors
as.data.frame(vif(overall_final_model))
# write the model to disk
saveRDS(object = overall_final_model, file = '../models/overall_final_model.RDS')
```

\newpage

### Televote

```{r model-televote-competition, echo = FALSE}
# fit step wise linear model
televote_competition_model <- step_lm_model(dataset = televote_data, pred_cols = competition_factors)
# extract out significant predictors
televote_sign_competition_factors <- televote_competition_model$sign_factors
```

```{r model-televote-performance, echo = FALSE}
# fit step wise linear model
televote_performance_model <- step_lm_model(dataset = televote_data, pred_cols = performance_factors)
# extract out significant predictors
televote_sign_performance_factors <- televote_performance_model$sign_factors
```

```{r model-televote-external, echo = FALSE}
# fit step wise linear model
televote_external_model <- step_lm_model(dataset = televote_data, pred_cols = external_factors)
# extract out significant predictors
televote_sign_external_factors <- televote_external_model$sign_factors
```

```{r model-televote-all, echo = FALSE}
# define the significant factors
televote_all_factors <- c(televote_sign_competition_factors, televote_sign_performance_factors, televote_sign_external_factors)
# fit step wise linear model
televote_all_model <- step_lm_model(dataset = televote_data, pred_cols = televote_all_factors)
# extract out significant predictors
televote_sign_all_factors <- televote_all_model$sign_factors
```

```{r model-televote-final, echo = FALSE}
# create a model formula
televote_final_model_form <- as.formula(paste('Points ~', paste(televote_sign_all_factors, collapse = ' + ')))
# create the maximum model
televote_final_model <- lm(televote_final_model_form, data = televote_data)
# get model summary
summary(televote_final_model)
# check variance inflation factors
as.data.frame(vif(televote_final_model))
# write the model to disk
saveRDS(object = televote_final_model, file = '../models/televote_final_model.RDS')
```

\newpage

### Jury 

```{r model-jury-competition, echo = FALSE}
# fit step wise linear model
jury_competition_model <- step_lm_model(dataset = jury_data, pred_cols = competition_factors)
# extract out significant predictors
jury_sign_competition_factors <- jury_competition_model$sign_factors
```

```{r model-jury-performance, echo = FALSE}
# fit step wise linear model
jury_performance_model <- step_lm_model(dataset = jury_data, pred_cols = performance_factors)
# extract out significant predictors
jury_sign_performance_factors <- jury_performance_model$sign_factors
```

```{r model-jury-external, echo = FALSE}
# fit step wise linear model
jury_external_model <- step_lm_model(dataset = jury_data, pred_cols = external_factors)
# extract out significant predictors
jury_sign_external_factors <- jury_external_model$sign_factors
```

```{r model-jury-all, echo = FALSE}
# define the significant factors
jury_all_factors <- c(jury_sign_competition_factors, jury_sign_performance_factors, jury_sign_external_factors)
# fit step wise linear model
jury_all_model <- step_lm_model(dataset = jury_data, pred_cols = jury_all_factors)
# extract out significant predictors
jury_sign_all_factors <- jury_all_model$sign_factors
```

```{r model-jury-final, echo = FALSE}
# create a model formula
jury_final_model_form <- as.formula(paste('Points ~', paste(jury_sign_all_factors, collapse = ' + ')))
# create the maximum model
jury_final_model <- lm(jury_final_model_form, data = jury_data)
# get model summary
summary(jury_final_model)
# check variance inflation factors
as.data.frame(vif(jury_final_model))
# write the model to disk
saveRDS(object = jury_final_model, file = '../models/jury_final_model.RDS')
```
