---
title: "MSc-ESC"
output: 
  html_document:
    toc: true
    df_print: kable
editor_options: 
  chunk_output_type: inline
---

# Section 4 - Data Modeling OVERVIEW

A stratified analysis method will be implemented
1. Test Overall Data
Iteratively create model step wise, 
2. Split and Test Data by Voting Method
Can't stratify the data by each country due to a lack of data
Need a minimum of 10/20 observations per covariate for regression analysis
Split data by voting method, research televote shows more bias than the jury
NOTE forward and step wise fitting will be utilized using AIC to determine model of best fit

NOTE: the models will be evaluated using the car package
MLR requires the residuals to be ~ IID N(0, sigma^2)
the residuals will be standardized for the assessment
Normality Assumptions will be accessed using:
(i) Normality tests from the nortest package
(ii) Visualizations such as histograms, QQ-plots, Residual Plots and Add Variable Plots
Constant Variance will be accessed using:
(i) non-constant variance test
Multi-collinearity will be accessed using:
(i) variance inflation factors
Outliers will be accessed using:
(i) Cooks Distance

```{r preliminaries}
# load the relevant libraries
library(rmarkdown)
library(knitr)
library(car)
library(dplyr)
# load custom functions
source('scripts/utilities/step_lm_model.R')
```

```{r data-load}
#-- Data --#

# load in the historic voting data for deriving the voting blocs
processed_data <- read.csv(file = "./data/processed_data.csv", header = T)
# split the televote data
televote_data <- processed_data %>% filter(Voting_Method_J == 0)
# split out the jury vote data
jury_data <- processed_data %>% filter(Voting_Method_J == 1)

# define the competition factors
competition_factors <- extract_preds_by_cats(cat = 'competition')
# define the performance factors
performance_factors <- extract_preds_by_cats(cat = 'performance')
# define the external factors
external_factors <- extract_preds_by_cats(cat = 'external')
```

## Model - Overall Data

iteratively create my own mode; 
continuously noting the magnitude and significance of the predictors. 
Fit the variables in groups/blocs; 
performance, competition and external variables. 
Fit each variable block independently. 
Then fit the all the best variables. 
Incorporate interaction and squared terms. 
Compare my final model with the baseline mode; 
making notes of any differences/similarities

```{r model-overall-competition}
# fit step wise linear model
overall_competition_model <- step_lm_model(dataset = processed_data, pred_cols = competition_factors)
# extract out significant predictors
overall_sign_competition_factors <- overall_competition_model$sign_factors
```

```{r model-overall-performance}
# fit step wise linear model
overall_performance_model <- step_lm_model(dataset = processed_data, pred_cols = performance_factors)
# extract out significant predictors
overall_sign_performance_factors <- overall_performance_model$sign_factors
```

```{r model-overall-external}
# fit step wise linear model
overall_external_model <- step_lm_model(dataset = processed_data, pred_cols = external_factors)
# extract out significant predictors
overall_sign_external_factors <- overall_external_model$sign_factors
```

```{r model-overall-all}
# define the significant factors
overall_all_factors <- c(overall_sign_competition_factors, overall_sign_performance_factors, overall_sign_external_factors)
# fit step wise linear model
overall_all_model <- step_lm_model(dataset = processed_data, pred_cols = overall_all_factors)
# extract out significant predictors
overall_sign_all_factors <- overall_all_model$sign_factors
```

```{r model-overall-final}
# create a model formula
overall_final_model_form <- as.formula(paste('Points ~', paste(overall_sign_all_factors, collapse = ' + ')))
# create the maximum model
overall_final_model <- lm(overall_final_model_form, data = processed_data)
# get model summary
summary(overall_final_model)
# get anova results
anova(overall_final_model)
# check variance inflation factors
vif(overall_final_model)
# write the model to disk
#saveRDS(object = overall_final_model, file = './models/overall_final_model.RDS')
```

## Model - Split by Televote

Consider how the televoting data differs from the jury data
with regards to:
performance features
competition features
external features

```{r model-televote-competition}
# fit step wise linear model
televote_competition_model <- step_lm_model(dataset = televote_data, pred_cols = competition_factors)
# extract out significant predictors
televote_sign_competition_factors <- televote_competition_model$sign_factors
```

```{r model-televote-performance}
# fit step wise linear model
televote_performance_model <- step_lm_model(dataset = televote_data, pred_cols = performance_factors)
# extract out significant predictors
televote_sign_performance_factors <- televote_performance_model$sign_factors
```

```{r model-televote-external}
# fit step wise linear model
televote_external_model <- step_lm_model(dataset = televote_data, pred_cols = external_factors)
# extract out significant predictors
televote_sign_external_factors <- televote_external_model$sign_factors
```

```{r model-televote-all}
# define the significant factors
televote_all_factors <- c(televote_sign_competition_factors, televote_sign_performance_factors, televote_sign_external_factors)
# fit step wise linear model
televote_all_model <- step_lm_model(dataset = televote_data, pred_cols = televote_all_factors)
# extract out significant predictors
televote_sign_all_factors <- televote_all_model$sign_factors
```

```{r model-televote-final}
# create a model formula
televote_final_model_form <- as.formula(paste('Points ~', paste(televote_sign_all_factors, collapse = ' + ')))
# create the maximum model
televote_final_model <- lm(televote_final_model_form, data = televote_data)
# get model summary
summary(televote_final_model)
# get anova results
anova(televote_final_model)
# check variance inflation factors
vif(televote_final_model)
# write the model to disk
#saveRDS(object = televote_final_model, file = './models/televote_final_model.RDS')
```

## Model - Split by Jury 

```{r model-jury-competition}
# fit step wise linear model
jury_competition_model <- step_lm_model(dataset = jury_data, pred_cols = competition_factors)
# extract out significant predictors
jury_sign_competition_factors <- jury_competition_model$sign_factors
```

```{r model-jury-performance}
# fit step wise linear model
jury_performance_model <- step_lm_model(dataset = jury_data, pred_cols = performance_factors)
# extract out significant predictors
jury_sign_performance_factors <- jury_performance_model$sign_factors
```

```{r model-jury-external}
# fit step wise linear model
jury_external_model <- step_lm_model(dataset = jury_data, pred_cols = external_factors)
# extract out significant predictors
jury_sign_external_factors <- jury_external_model$sign_factors
```

```{r model-jury-all}
# define the significant factors
jury_all_factors <- c(jury_sign_competition_factors, jury_sign_performance_factors, jury_sign_external_factors)
# fit step wise linear model
jury_all_model <- step_lm_model(dataset = jury_data, pred_cols = jury_all_factors)
# extract out significant predictors
jury_sign_all_factors <- jury_all_model$sign_factors
```

```{r model-jury-final}
# create a model formula
jury_final_model_form <- as.formula(paste('Points ~', paste(jury_sign_all_factors, collapse = ' + ')))
# create the maximum model
jury_final_model <- lm(jury_final_model_form, data = jury_data)
# get model summary
summary(jury_final_model)
# get anova results
anova(jury_final_model)
# check variance inflation factors
vif(jury_final_model)
# write the model to disk
#saveRDS(object = jury_final_model, file = './models/jury_final_model.RDS')
```
