---
title: "MSc-ESC"
output: 
  html_document:
    toc: true
    df_print: kable
editor_options: 
  chunk_output_type: inline
---

## Section 6 - Conclusions

This script specializes in analyzing the models and drawing inferences for the research question. The research question is whether voting blocs, echo nest music features and migration patterns can explain the points and voting patterns of the 2016 ESC. This can be converted into a statistical problem using multiple linear regression whereby determining whether voting blocs, echo nest music features and migration patterns significantly explain the points and voting patterns.

This question will be answer by:
1. Constricting t-tests
2. Observing the signs of the estimated coefficients
3. measuring the increase in variance explained (R-sq) with the addition of a predictor variable

```{r preliminaries}
# load relevant libraries
library(rmarkdown)
library(knitr)
library(car)
library(dplyr)
```

```{r data-load}
# load in the historic voting data for deriving the voting blocs
processed_data <- read.csv(file = "./data/arch/processed_data.csv", header = T)
# split the televote data
televote_data <- processed_data %>% filter(Voting_Method_J == 0)
# split out the jury vote data
jury_data <- processed_data %>% filter(Voting_Method_J == 1)
```

## Overall Model

```{r overall-model}
#-- Fit the Final Model --#

# define the model formula for the final model
my_model_overall_form <- Points ~ Average_Points + VBlocs1_TC_3 + CAP_DIST_km + 
                                  FC_NonCitzens + ComLANGFAM_y + liveness + key_3 + METRIC_Citizens + 
                                  TC_PerfType_Solo + key_2 + VBlocs1_TC_13 + key_6 + time_signature_4 + 
                                  ComVBlocs1_y + VBlocs1_TC_1 + key_5 + OOA + speechiness
# fit the final overall model
my_model_overall <- lm(formula = my_model_overall_form, data = processed_data)
# perform box-cox transformation
bct <- boxCox(object = my_model_overall)
# return the optimal power transformation
p <- bct$x[which.max(x = bct$y)]
# transform the response variable
bctPoints <- (((processed_data$Points)^p) - 1)/(p)
# redefine the model formula for the final model
my_model_overall_bct_from <- bctPoints ~ Average_Points + VBlocs1_TC_3 + CAP_DIST_km + 
                                          FC_NonCitzens + ComLANGFAM_y + liveness + key_3 + METRIC_Citizens + 
                                          TC_PerfType_Solo + key_2 + VBlocs1_TC_13 + key_6 + time_signature_4 + 
                                          ComVBlocs1_y + VBlocs1_TC_1 + key_5 + OOA + speechiness
# refit the model with the power transformation
my_model_overall <- lm(my_model_overall_bct_from, data = processed_data)
# generate final model summary
summary(my_model_overall)
# write the model to disk
saveRDS(object = my_model_overall, file = './models/arch/overall_final_model.RDS')
```

## T-tests

Make sure to report 
1. the adjusted R-sq value
2. the degrees of freedom
3. the f statistic for f test
4. the p-value for f test

NOTE: include notation for levels of significance 
* for 0.05
** for 0.01
*** for 0.001

```{r overall-t-tests}
summary(my_model_overall)
```

## Directional / Sign effects

Report the sign effects of each significant coefficient
+ indicates the predictor variable has a positive effect on the dependent variable
- indicates the predictor variables  has a negative effect on the dependent variable

```{r overall-sign-effects}
summary(my_model_overall)
```

## Explained Variance

Investigate the increase of variance explained by incorporating specific predictor variables. Observe the increase in R-sq when a predictor variable is included / excluded from the model. Do this for voting blocs, Echo Nest music factors and Migration patterns

```{r overall-variance-vb}
# define the model formula
omodel_ex.vblocs_form <- bctPoints ~ Average_Points + CAP_DIST_km + 
                                     FC_NonCitzens + ComLANGFAM_y + liveness + key_3 + METRIC_Citizens + 
                                     TC_PerfType_Solo + key_2 + key_6 + time_signature_4 + 
                                    key_5 + OOA + speechiness
# fit the linear model
omodel_ex.vblocs <- lm(formula = omodel_ex.vblocs_form, data = processed_data)
# generate model summary
summary(omodel_ex.vblocs)
```

```{r overall-variance-en}
# define the model formula
omodel_ex.music_form <- bctPoints ~ Average_Points + VBlocs1_TC_3 + CAP_DIST_km + 
                                    FC_NonCitzens + ComLANGFAM_y + METRIC_Citizens + 
                                    TC_PerfType_Solo + VBlocs1_TC_13 + 
                                     ComVBlocs1_y + VBlocs1_TC_1 + OOA
# fit the linear model
omodel_ex.music <- lm(omodel_ex.music_form, data = processed_data)
# investigate the R-sq value
summary(omodel_ex.music)
```

```{r overall-variance-migration}
# define the model formula
omodel_ex.mig_form <- Points ~ Average_Points + VBlocs1_TC_3 + CAP_DIST_km + 
                               ComLANGFAM_y + liveness + key_3 + 
                               TC_PerfType_Solo + key_2 + VBlocs1_TC_13 + key_6 + time_signature_4 + 
                               ComVBlocs1_y + VBlocs1_TC_1 + key_5 + OOA + speechiness
# fit the linear model
omodel_ex.mig <- lm(formula = omodel_ex.mig_form, data = processed_data)
# generate the model summary
summary(omodel_ex.mig)
```

## TELEVOTE MODEL

```{r televote-model}
# define the model formula
my_model_tele_form <- Points ~ METRIC_Citizens + Average_Points + VBlocs1_TC_3 + 
                               VBlocs2_TC_1 + mode_1 + key_11 + OOA + acousticness + danceability + 
                               key_7 + VBlocs1_TC_13 + ComLANGFAM_y
# fit the linear model
my_model_tele <- lm(formula = my_model_tele_form, data = televote_data)
# perform box-cox transformation
bct <- boxCox(object = my_model_tele)
# return the optimal power transformation
p <- bct$x[which.max(x = bct$y)]
# transform the response variable
bctPoints <- (((televote_data$Points)^p) - 1)/(p)
# define the model formula 
my_model_tele_bct <- bctPoints ~ METRIC_Citizens + Average_Points + VBlocs1_TC_3 + 
                                 VBlocs2_TC_1 + mode_1 + key_11 + OOA + acousticness + danceability + 
                                 key_7 + VBlocs1_TC_13 + ComLANGFAM_y
# refit the model with the power transformation
my_model_tele <- lm(my_model_tele_bct, data = televote_data)
# generate model summary
summary(my_model_tele)
# write the model to disk
saveRDS(object = my_model_tele, file = './models/arch/televote_final_model.RDS')
```

## T-tests

Make sure to report 
1. the adjusted R-sq value
2. the degrees of freedom
3. the f statistic for f test
4. the p-value for f test

NOTE: include notation for levels of significance 
* for 0.05
** for 0.01
*** for 0.001

```{r televote-t-tests}
summary(my_model_tele)
```

## Directional / Sign effects 

Report the sign effects of each significant coefficient
+ indicates the predictor variable has a positive effect on the dependent variable
- indicates the predictor variables  has a negative effect on the defendant variable

```{r televote-sign-effects}
summary(my_model_tele)
```

## Explained Variance

Investigate the increase of variance explained by incorporating specific predictor variables. Observe the increase in R-sq when a predictor variable is included / excluded from the model. Do this for voting blocs, Echo Nest music factors and Migration patterns

```{r televote-variance-vb}
# define model formula
tmodel_ex.vblocs_form <- bctPoints ~ METRIC_Citizens + Average_Points + 
                                     mode_1 + key_11 + OOA + acousticness + danceability + 
                                     key_7 + ComLANGFAM_y
# fit linear model
tmodel_ex.vblocs <- lm(tmodel_ex.vblocs_form, data = televote_data)
# investigate the R-sq value
summary(tmodel_ex.vblocs)
```

```{r televote-variance-en}
# define model formula
tmodel_ex.music_form <- bctPoints ~ METRIC_Citizens + Average_Points + VBlocs1_TC_3 + 
                                    VBlocs2_TC_1 + OOA + VBlocs1_TC_13 + ComLANGFAM_y
# fit linear model
tmodel_ex.music <- lm(tmodel_ex.music_form, data = televote_data)
# investigate the R-sq value
summary(tmodel_ex.music)
```

```{r televote-variance-migration}
# define model formula
tmodel_ex.mig_form <- bctPoints ~  Average_Points + VBlocs1_TC_3 + 
                                   VBlocs2_TC_1 + mode_1 + key_11 + OOA + acousticness + danceability + 
                                   key_7 + VBlocs1_TC_13 + ComLANGFAM_y
# fit linear model
tmodel_ex.mig <- lm(tmodel_ex.mig_form, data = televote_data)
# investigate the R-sq value
summary(tmodel_ex.mig)
```

## JURY MODEL

```{r jury-model}
# power transformation of 3/4
# define model formula
# Note: thesis shows output for non-power transformed case (i.e. Points ~ VBlocs2_TC_4 + key_3 + ... )
my_model_jury_form <- (Points)^(3/4) ~ VBlocs2_TC_4 + key_3 + TC_PerfType_Solo + 
                                       liveness + ComVBlocs1_y + ComLANGFAM_y
# fit linear model
my_model_jury <- lm(formula = my_model_jury_form, data = jury_data)
# generate model summary
summary(my_model_jury)
# write the model to disk
saveRDS(object = my_model_jury, file = './models/arch/jury_final_model.RDS')
```

## T-tests

Make sure to report 
1. the adjusted R-sq value
2. the degrees of freedom
3. the f statistic for f test
4. the p-value for f test

NOTE: include notation for levels of significance 
* for 0.05
** for 0.01
*** for 0.001

```{r jury-t-tests}
summary(my_model_jury)
```

## Directional / Sign effects

Report the sign effects of each significant coefficient
+ indicates the predictor variable has a positive effect on the dependent variable
- indicates the predictor variables  has a negative effect on the dependent variable

```{r jury-sign-effects}
summary(my_model_jury)
```

## Explained Variance

Investigate the increase of variance explained by incorporating specific predictor variables. Observe the increase in R-sq when a predictor variable is included / excluded from the model. Do this for voting blocs, Echo Nest music factors and Migration patterns

```{r jury-variance-vb}
# define model formula
jmodel_ex.vblocs_form <- (Points)^(3/4) ~ key_3 + TC_PerfType_Solo + 
                                          liveness + ComVBlocs1_y + ComLANGFAM_y
# fit linear model
jmodel_ex.vblocs <- lm(formula = jmodel_ex.vblocs_form, data = jury_data)
# investigate the R-sq value
summary(jmodel_ex.vblocs)
```

```{r jury-en}
# define model formula
jmodel_ex.music_form <- (Points)^(3/4) ~ VBlocs2_TC_4 + TC_PerfType_Solo + 
                                         ComVBlocs1_y + ComLANGFAM_y
# fit linear model
jmodel_ex.music <- lm(formula = jmodel_ex.music_form, data = jury_data)
# investigate the R-sq value
summary(jmodel_ex.music)
```

